Pre-processing high frequency animal tracking data

### 2020-10-06 03_smoothing_tracks.Rmd
some more changes

# now prep gap
fig_smooth_1d <- 
  ggplot()+
  geom_path(data = data[1:2000,],
            aes(time, y),
            size = 0.2)+
  geom_point(data = rbindlist(data_plot),
             aes(time, y,
                 col = factor(window_size)),
             alpha = 0.5,
             size = 1,
             show.legend = FALSE,
             shape = 16)+
  coord_cartesian(xlim = c(5000, 7000),
                  ylim = c(0.7, NA),
                  expand = F)+
  scale_colour_brewer(palette = "Spectral")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank())+
  labs(x = "position index", y = "Y coordinate")
```

@@

# make basic plot with rectangle
fig_bounds <- 
  ggplot()+
  geom_path(data = data,
            aes(x, y))+
  geom_hline(yintercept = c(0.6, 0.8),
             col = "grey20",
             lty = 2, lwd = 0.2)+
  coord_equal(expand = F) +
  ggthemes::theme_map()

>>>>

# now prep gap
fig_smooth_1d <-
  ggplot()+
  geom_point(data = rbindlist(data_plot),
             aes(time, y,
                 col = factor(window_size)),
             alpha = 0.5,
             size = 1,
             show.legend = FALSE,
             shape = 16)+
  coord_cartesian(xlim = c(5000, 7000),
                  ylim = c(0.7, NA),
                  expand = F)+
  scale_colour_brewer(palette = "Spectral")+
  ggthemes::theme_few()+
  theme(axis.text = element_blank())+
  labs(x = "position index", y = "Y coordinate")
```

@@

# make basic plot with rectangle
fig_bounds <-
  ggplot()+
  geom_point(data = data_errors,
            aes(x, y),
            size = 0.1, alpha = 0.2, 
            col = "grey")+
  geom_path(data = data,
            aes(x, y),
            size = 0.1)+
  geom_hline(yintercept = c(0.6, 0.8),
             col = "grey20",
             lty = 2, lwd = 0.2)+
  coord_equal(expand = F) +
  ggthemes::theme_map()

### 2020-10-06 05_calibration_data.Rmd
some more changes

```{r}
# load libs
library(data.table)
library(atlastools)
library(ggplot2)
```

@@

```{r}
# plot data
fig_data_bbox <-
  ggplot()+
  geom_path(data = data_unproc,
            aes(x, y),
             col = "grey", alpha = 0.5, size = 0.3)+
  geom_point(data = data_unproc,
             aes(x, y,
                 col = x > 645000),
             alpha = ifelse(data_unproc$x > 645000, 0.5, 1),
             size = ifelse(data_unproc$x > 645000, 0.3, 2),
             show.legend = F)+
  scale_color_brewer(palette = "Set1")+
  geom_vline(xintercept = 645000,
             col = "grey", lty = 2)+
  ggspatial::annotation_scale(location = "br")+
  ggthemes::theme_map()+
  coord_sf(crs = 32631)

@@

```{r}
# make figure
fig_smooth <-
  ggplot()+
  geom_path(data = data,
             aes(x, y),
            col = RColorBrewer::brewer.pal(4, "Set1")[4])+
  annotate(geom = "rect",
           xmin = c(650785, 650045),
           ymin = c(5904450, 5902700),
           xmax = c(652500, 650682),
           ymax = c(5906133, 5903080),
           fill = "grey90",
           alpha = 0.5,
           col = NA)+
    annotate(geom = "text",
           x = c(650785, 650045) + 400,
           y = c(5904450, 5902700) + 300,
           col = "black",
           label = c("(a)", "(b)"))+
  theme_void()+
  coord_sf(crs = 32631)+
  labs(colour = "track")

# make zoomed in figures
fig_inset <- mapply(function(xl, yl) {
  ggplot()+
    geom_point(data = data_raw[!data_unproc, on = c("x", "y")],
               aes(x, y),
               col = "grey",
               shape = 4,
               size = 0.4)+
    geom_point(data = data_unproc,
               aes(x, y),
               col = RColorBrewer::brewer.pal(3, "Set1")[3],
               shape = 1, 
               alpha = 0.5)+
    geom_path(data = data,
              aes(x, y),
              col = RColorBrewer::brewer.pal(4, "Set1")[4])+
    ggthemes::theme_few()+
    theme(axis.title = element_blank(),
          axis.text = element_blank())+
    ggspatial::annotation_scale(location = "br")+
    coord_sf(crs = 32631,
             xlim = xl,
             ylim = yl, expand = F
             )+
    labs(colour = "track")
}, list(xl1 = c(650785, 652500),
        xl2 = c(650045, 650682)),
   list(yl1 = c(5904450, 5906133),
        yl2 = c(5902700, 5903080)),
SIMPLIFY = FALSE)

@@

wrap_plots(fig_inset)+
  plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")")
# save figure
ggsave(filename = "figures/fig_calib_median_smooth.png",
       width = 185 / 25, height = 90 / 25)
```

@@

# make a copy of only residence points
data_res <- data[residence == T, ]

# get a simple inverse of speed as residence time classification
data_slow <- copy(data_unproc)
data_slow <- data_slow[speed_in < 1, ]
```

@@

```{r}
# on known residence points
patch_res_known <- atl_res_patch(data_res, 
                                buffer_radius = 10,
                                lim_spat_indep = 100,
                                lim_time_indep = 30,
                                min_fixes = 3)

# on data where the residence is unknown
patch_res_unknown <- atl_res_patch(data_slow, 
                                buffer_radius = 10,
                                lim_spat_indep = 100,
                                lim_time_indep = 30,
                                min_fixes = 10)
```

@@

```{r}
# for the known and unkniwn patches
patch_sf_data <- lapply(list(patch_res_known, patch_res_unknown),
                        atl_patch_summary, which_data = "spatial")

# assign a crs
patch_sf_data <- lapply(patch_sf_data, function(df) {
  sf::st_crs(df) <- 32631
  return(df)
})

patch_summary_data <- lapply(list(patch_res_known, patch_res_unknown),
                        atl_patch_summary)
```

@@

```{r}
# patch with residence points and walking patches
fig_basic_residence <-
  ggplot()+
  geom_point(data = data,
             aes(x, y,
                 col = residence),
            size = ifelse(data$residence, 4, 0.2),
            shape = ifelse(data$residence, 2, 1),
            show.legend = FALSE)+
  scale_color_brewer(palette = "Set1")+
  ggthemes::theme_map()+
  coord_sf(crs = 32631)

# figure of residence patches known
# fig_patch_known <- 
  ggplot()+
  geom_sf(data = patch_sf_data[[1]],
          fill = "steelblue", col = NA)+
  geom_path(data = patch_summary_data[[1]],
            aes(x_median, y_median),
            col = "red")+
  ggthemes::theme_map()

pla = mapply(function(pol) {

  bbox = st_bbox(pol)

  ggplot()+
    geom_path(data = data_unproc,
               aes(x, y),
               col = RColorBrewer::brewer.pal(4, "Set1")[4])+
    geom_sf(data = pol,
            alpha = 0.2)+
    ggthemes::theme_map()+
    coord_sf(
      expand = F,
      xlim = c(bbox[["xmin"]], bbox[["xmax"]]),
      ylim = c(bbox[["ymin"]], bbox[["ymax"]]))
}, patch_sf_data[[1]]$polygons, SIMPLIFY = F)

fig_patches_small <- wrap_plots(pla)+
  plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")")

# figure of unknown patches
fig_patch_unknown <-
  ggplot()+
  geom_path(data = patch_summary_data[[1]],
            aes(x_median, y_median),
            col = "black", size = 0.2, lty = 2)+
  geom_sf(data = patch_sf_data[[2]],
          fill = "steelblue", col = NA)+
  geom_segment(data = patch_summary_data[[2]],
               aes(x = x_end,
                   xend = shift(x_start, type = "lead"),
                   y = y_end,
                   yend = shift(y_start, type = "lead")),
               col = "red"
  )+
  ggthemes::theme_map()

# arrange plots
fig_res_patch <- 
  wrap_plots(list(fig_basic_residence, fig_patch_known, fig_patch_unknown),
           design = "AAAA\nAAAA\nBBCC\nBBCC")+
  plot_annotation(tag_levels = "a", 
                  tag_prefix = "(", tag_suffix = ")")

# save figure
ggsave(fig_res_patch, filename = "figures/fig_calib_respatch.png",
       height = 200/25, width = 185/25)
```

@@

# round median coordinate for inferred patches
patch_summary_inferred <- 
        patch_summary_data[[1]][, 
                                c("x_median", "y_median", 
                                  "nfixes", "duration")
                                ][, `:=`(x_median = round(x_median, digits = -2),
                                         y_median = round(y_median, digits = -2))]

@@

### Plot known and unknown patches

```{r}
# make figure comparing different methods
ggplot()+
  geom_point(data = patch_summary_compare,
             aes(duration_real, 
                 duration),
             show.legend = F,
            size = 4,
            shape = 1)+
  geom_smooth(data = patch_summary_compare,
             aes(duration_real, 
                 duration),
             se = F,
             method = "glm",
             show.legend = F)+
  # scale_x_log10()+
  # scale_y_log10()+
  # coord_cartesian(xlim = c(10, NA))+
  ggthemes::theme_few()+
  labs(x = "# positions (real)",
       y = "# positions (inferred)")
```

>>>>

```{r}
# load libs
library(data.table)
library(atlastools)
library(ggplot2)
library(patchwork)

# prepare a palette
pal <- RColorBrewer::brewer.pal(4, "Set1")
```

@@

```{r}
# plot data
fig_data_bbox <-
  ggplot()+
  geom_path(data = data_raw,
            aes(x, y),
             col = "grey", alpha = 0.5, size = 0.3)+
  geom_point(data = data_raw,
             aes(x, y,
                 col = x > 645000),
             alpha = ifelse(data_raw$x > 645000, 0.5, 1),
             size = ifelse(data_raw$x > 645000, 0.3, 2),
             show.legend = F)+
  scale_color_brewer(palette = "Set1")+
  geom_vline(xintercept = 645000,
             col = "grey", lty = 2)+
  ggspatial::annotation_scale(location = "br")+
  ggthemes::theme_map()+
  coord_sf(crs = 32631)

@@

```{r}
# make figure
fig_smooth <-
  ggplot()+
  geom_path(data = data,
             aes(x, y),
            col = RColorBrewer::brewer.pal(4, "Set1")[4])+
  annotate(geom = "rect",
           xmin = c(650785, 650045),
           ymin = c(5904450, 5902700),
           xmax = c(652500, 650682),
           ymax = c(5906133, 5903080),
           fill = "grey90",
           alpha = 0.5,
           col = NA)+
    annotate(geom = "text",
           x = c(650785, 650045) + 400,
           y = c(5904450, 5902700) + 300,
           col = "black",
           label = c("d", "e"))+
  theme_void()+
  coord_sf(crs = 32631)+
  labs(colour = "track")

# make zoomed in figures
fig_inset <- mapply(function(xl, yl) {
  ggplot()+
    geom_point(data = data_raw[!data_unproc, on = c("x", "y")],
               aes(x, y),
               col = "grey",
               shape = 4,
               size = 0.4)+
    geom_point(data = data_unproc,
               aes(x, y),
               col = RColorBrewer::brewer.pal(3, "Set1")[3],
               shape = 1, 
               alpha = 0.5)+
    geom_path(data = data,
              aes(x, y),
              col = RColorBrewer::brewer.pal(4, "Set1")[4])+
    ggthemes::theme_few()+
    theme(axis.title = element_blank(),
          axis.text = element_blank())+
    ggspatial::annotation_scale(location = "br")+
    coord_cartesian(
             xlim = xl,
             ylim = yl, expand = F
             )+
    labs(colour = "track")
}, list(xl1 = c(650785, 652500),
        xl2 = c(650045, 650682)),
   list(yl1 = c(5904450, 5906133),
        yl2 = c(5902700, 5903080)),
SIMPLIFY = FALSE)

@@

fig_median_smooth <- 
  wrap_plots(fig_inset)+
  plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")")
# save figure
ggsave(filename = "figures/fig_calib_median_smooth.png",
       width = 185 / 25, height = 90 / 25)
```

### Plot pre-processing steps

```{r}
# make combined walkthrough figure
figure_walkthrough <- 
  wrap_plots(
  append(
    list(
      fig_data_bbox, fig_low_nbs, fig_speed_outliers), 
    fig_inset),
design = "ABD\nACE")+
  plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")")

# save combined figure
ggsave(figure_walkthrough,
       filename = "figures/fig_walkthrough.png",
       height = 130, width = 225, units = "mm")
```

@@

# make a copy of only residence points
data_res <- data[residence == T, ]
```

@@

```{r}
# on known residence points
patch_res_known <- atl_res_patch(data_res, 
                                buffer_radius = 10,
                                lim_spat_indep = 100,
                                lim_time_indep = 30,
                                min_fixes = 3)
```

@@

```{r}
# for the known and unkniwn patches
patch_sf_data <- atl_patch_summary(patch_res_known, which_data = "spatial")

# assign crs
sf::st_crs(patch_sf_data) <- 32631

# get summary data
patch_summary_data <- atl_patch_summary(patch_res_known, which_data = "summary")
```

@@

```{r}
# patch with residence points and all patches
fig_basic_residence <-
  ggplot()+
  geom_path(data = data,
            aes(x, y),
            lwd = 0.2,
            col = "grey20")+
  geom_point(data = data_res,
             aes(x, y),
             shape = 2,
             size = 2,
             col = pal[2])+
  geom_path(data = patch_summary_data,
            aes(x_median + 200, y_median),
            col = "orange", lty = 1)+
  geom_point(data = patch_summary_data,
             aes(x_median + 200, y_median,
                 fill = duration / 60),
             size = 5,
             shape = 21)+

  scale_fill_distiller(palette = "YlOrRd",
                       direction = 1,
                       trans = "log10",
                       limits = c(1, 60))+
  ggthemes::theme_map()+
  theme(legend.position = c(0, 0.3),
        legend.key.width = unit(3, units = "mm"),
        legend.background = element_blank())+
  ggspatial::annotation_scale(location = "tl")+
  coord_sf(crs = 32631)+
  labs(fill = "inferred\nduration (min)")
```

@@

# round median coordinate for inferred patches
patch_summary_inferred <- 
  patch_summary_data[, 
                     c("x_median", "y_median", 
                       "nfixes", "duration")
                     ][, `:=`(x_median = round(x_median, digits = -2),
                              y_median = round(y_median, digits = -2))]

@@

### Plot durations comparisons

```{r}
# get linear model
model_duration <- lm(duration ~ duration_real,
                     data = patch_summary_compare)

# get R2
summary(model_duration)
```

```{r}
# make figure comparing different methods
figure_patch_duration <-
  ggplot()+
  geom_segment(data = patch_summary_compare,
               aes(x = 100,
                   xend = duration_real,
                   y = duration,
                   yend = duration),
               col = "grey")+
  geom_segment(data = patch_summary_compare,
               aes(x = duration_real,
                   xend = duration_real,
                   y = 100,
                   yend = duration),
               col = "grey")+
  geom_point(data = patch_summary_compare,
             aes(duration_real, 
                 duration),
             show.legend = F,
            size = 3,
            stroke = 1,
            shape = 21,
            fill = pal[2],
            col = "white")+
  geom_smooth(data = patch_summary_compare,
             aes(duration_real, 
                 duration),
             se = F,
             size = 0.5,
             col = pal[1],
             alpha = 0.2,
             method = "glm",
             show.legend = F)+
  annotate(geom = "text",
           x = 300, y = 900,
           size = 5,
           label = "R^2 == 0.845",
           parse = T)+
  scale_x_log10(breaks = c(300, 600, 900),
                labels = as.integer(c(300, 600, 900) / 60))+
  scale_y_log10(breaks = c(300, 600, 900),
                labels = as.integer(c(300, 600, 900) / 60))+
  coord_equal(expand = F,
              xlim = c(200, 1200),
              ylim = c(200, 1200))+
  ggthemes::theme_few()+
  labs(x = "duration (min)",
       y = "inferred duration (min)")
```

### Join patch map and duration comparison figure

```{r}
# wrap together
figure_res_patch <-
  wrap_plots(list(fig_basic_residence, figure_patch_duration))+
    plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")") &
  theme(plot.tag = element_text(face = "bold"))

# save figure
ggsave(figure_res_patch,
       filename = "figures/fig_calib_residence_patch.png",
       height = 130, width = 225, units = "mm")
```
